---
title: "vitoria-project"
author: "vitoria"
date: "2025-04-02"
output: html_document
---

# Load libraries

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(httr2)
library(jsonlite)
library(ggplot2)
library(readxl)
library(rvest)
library(readr)
library(stringr)
library(tidyr)
library(scales)
library(httr)
library(xml2)
```

# Functions to get SMILES from CID

```{r message=FALSE, warning=FALSE}
# Try to get CID from compound name
# If not possible, try to get CID from SID
# Param - name: compound name (character)
# Return - compound_cid: CID (integer)
get_cid <- function(name) {
  if (is.na(name) || name == "") return(NA)  # Handle missing names
  
  encoded_name <- URLencode(name, reserved = TRUE)
  
  # Try getting CID using compound name endpoint
  url_cid <- paste0("https://pubchem.ncbi.nlm.nih.gov/rest/pug/compound/name/", 
                    encoded_name, "/cids/JSON")
  response_cid <- tryCatch({
    req <- request(url_cid) |> req_perform()
    if (resp_status(req) == 200) {
      data <- resp_body_json(req)
      if (!is.null(data$IdentifierList$CID)) {
        compound_id <- data$IdentifierList$CID[[1]]
        returncompound_id  # Return first CID
      }
    }
    return(NA)
  }, error = function(e) return(NA))
  
  # If a CID was found, return it immediately
  if (!is.na(response_cid)) {
    return(response_cid)
  }
  
  # Pause to respect API rate limits
  Sys.sleep(0.5)  
  
  # Try getting SID using substance name endpoint
  url_sid <- paste0("https://pubchem.ncbi.nlm.nih.gov/rest/pug/substance/name/", encoded_name, "/sids/JSON")
  response_sid <- tryCatch({
    req <- request(url_sid) |> req_perform()
    if (resp_status(req) == 200) {
      data <- resp_body_json(req)
      if (!is.null(data$IdentifierList$SID) && length(data$IdentifierList$SID) > 0) {
        data$IdentifierList$SID[[1]]  # Store the SID 
      } else {
        NA
      }
    } else {
      NA
    }
  }, error = function(e) return(NA))
  
  # Pause to respect API rate limits
  Sys.sleep(0.5)  
  
  # Try getting CID using SID
  url_sid_cid <- paste0("https://pubchem.ncbi.nlm.nih.gov/rest/pug/substance/sid/", 
                        response_sid, "/JSON")
  response_cid <- tryCatch({
    req <- request(url_sid_cid) |> req_perform()
    if (resp_status(req) == 200) {
      data <- resp_body_json(req)
      if (!is.null(data$PC_Substances[[1]]$compound[[2]]$id$id$cid) && length(data$PC_Substances[[1]]$compound[[2]]$id$id$cid) > 0) {
        # print("Found CID using SID") 
        compound_id <- data$PC_Substances[[1]]$compound[[2]]$id$id$cid[[1]]
        return(compound_id)  # Return CID
      } 
    }
    return(NA)
  }, error = function(e) return(NA))
}

# Get SMILES from Compound ID
# Param - compound_id: CID (integer) 
# Return (character): SMILES for the compound
get_smiles_from_cid <- function(compound_id) {

  url <- paste0("https://pubchem.ncbi.nlm.nih.gov/rest/pug/compound/cid/", compound_id, "/property/CanonicalSMILES/JSON")
  
  response <- request(url) |> req_perform()
  
  # If there was a response, get the data
  if (resp_status(response) == 200) {
    data <- resp_body_json(response)
    if (!is.null(data$PropertyTable$Properties[[1]]$CanonicalSMILES)) {
      compound_smiles <- data$PropertyTable$Properties[[1]]$CanonicalSMILES
      return(compound_smiles)
    }
  }
  
  return(NA)
}
```

# Read and preprocess the data from gutMGene

```{r}
# Read the CSV file from gutMGene
gutmgene <- read.csv("data/gutmgene.csv", stringsAsFactors = FALSE)

# Rename columns to avoid special character issues
colnames(gutmgene) <- c("Host_Species", "Gut_Microbe_ID", "Rank",
                        "Metabolite_ID", "Evidence_Type", "Evidence_Amount")

cat("N metabolites raw file:",
    length(gutmgene$Metabolite_ID), "\n")

unique_species_count <- gutmgene %>% filter(Rank == 'genus') %>%
  summarise(count = n_distinct(Gut_Microbe_ID)) %>% pull(count)
cat("N microbes in raw file:", unique_species_count, "\n")

cat("N of unique genera:", sum(gutmgene$Rank == 'genus'), "\n")

# Extract the metabolite name from the "Metabolite_ID" column
gutmgene <- gutmgene %>%
  mutate(
    Metabolite_Name = str_remove(Metabolite_ID, "\\([^()]*\\)$"), # Removes last set of parentheses and content
    Metabolite_Name = str_trim(Metabolite_Name), # Trims whitespace
    Gut_Microbe_Name = str_extract(Gut_Microbe_ID, "^[^(]+"), # Extracts text before '('
    Gut_Microbe_ID = as.integer(str_extract(Gut_Microbe_ID, "(?<=\\().*(?=\\))")), # Extracts numbers inside '()'
  )

# Get CID with vectorized function
get_cid_vec <- Vectorize(get_cid)
gutmgene <- gutmgene %>%
  mutate(
    Metabolite_CID = get_cid_vec(Metabolite_Name) # Obtains CID for the metabolite name
  )
```

# Build plots of raw data

## Plot 1 - Distribution of genera among metabolites

```{r}
# Filter only genus-level microbes
gutmgene_filtered <- gutmgene[gutmgene$Rank == "genus", ]

# Count occurrences of each genus
genus_counts <- table(gutmgene_filtered$Gut_Microbe_Name)

# Filter to only include genera with frequency > 5
genus_to_keep <- names(genus_counts[genus_counts > 5])
gutmgene_filtered <- gutmgene_filtered[gutmgene_filtered$Gut_Microbe_Name %in% genus_to_keep, ]

# Order from most frequent to less frequent
gutmgene_filtered_count <- gutmgene_filtered %>%
  count(Gut_Microbe_Name) %>%
  mutate(Gut_Microbe_Name = reorder(Gut_Microbe_Name, -n))

# Create the plot
microbe_distribution <- ggplot(gutmgene_filtered_count,
                               aes(x = Gut_Microbe_Name, y = n)) +
  geom_bar(stat = "identity", fill = "gray30", width = 0.5) +
  theme_minimal() +  
  theme(
    axis.title.y = element_text(size = 16),
    axis.title.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16, angle = 60, hjust = 1),
    panel.grid = element_blank(),  # Remove grid lines
    axis.line.x = element_line(color = "black", linewidth = 0.5),  # Only X-axis line
    axis.line.y = element_line(color = "black", linewidth = 0.5),  # Only Y-axis line
    axis.ticks = element_line(color = "black", linewidth = 1)      # Ensure ticks are visible
  ) +
  labs(x = "Gut Microbe (Genus)", 
       y = "Count") +
  scale_y_continuous(breaks = seq(0,max(gutmgene_filtered_count$n),
                                  by = 5)) +                       # Sety axis scale
  coord_cartesian(ylim = c(0, max(gutmgene_filtered_count$n)))     # Set the grid's height

# Create a directory for plots if it doesn't exist
if (!dir.exists("plots")) {
  dir.create("plots")
}

# Save plot as PNG
png("plots/microbe_distribution.png", width = 350, height = 300)
print(microbe_distribution)
dev.off()
```

## Plot 2 - Distribution of how many evidences the metabolite-microbe rows have

```{r}
evidence_amount_counts <- gutmgene %>%
  count(Evidence_Amount)  # Count each unique value

# Create the plot
evidence_amount <- ggplot(evidence_amount_counts, aes(x = factor(Evidence_Amount), y = n)) +  
  geom_col(fill = "gray30", width = 0.5) +
  geom_text(aes(label = n), vjust = -0.5, size = 5, color = "black") +  # Add text labels
  theme_minimal() +  
  theme(
    axis.title.y = element_text(size = 16),
    axis.title.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16, hjust = 1),
    panel.grid = element_blank(),
    axis.line.x = element_line(color = "black", linewidth = 0.5),
    axis.line.y = element_line(color = "black", linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 1)
  ) +
  scale_y_continuous(breaks = pretty_breaks(n = 7)) +             # Set y axis scale
  labs(x = "Evidence Amount", 
       y = "Count") +  
  coord_cartesian(ylim = c(0, max(evidence_amount_counts)*1.03))  # Set the grid's height

# Save plot as PNG
png("plots/evidence_amount.png", width = 350, height = 300)
print(pevidence_amount)
dev.off()
```

## Plot 3 - Distribution of evidence types (causally or correlatively)

```{r}
# Split the 'Evidence' column by commas, then count occurrences of each type of evidence
evidence_count <- gutmgene %>%
  separate_rows(Evidence_Type, sep = ";") %>%  # Split 
  count(Evidence_Type) %>%  # Count
  filter(Evidence_Type %in% c("causally", "correlatively"))

# Create the plot
evidence_types <- ggplot(evidence_count, aes(x = "", y = n, fill = Evidence_Type)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +    # Pie chart
  theme_void() +                # No axes
  scale_fill_manual(values = c("causally" = "gray30", "correlatively" = "gray50")) +
  geom_text(aes(label = paste(n, "(", round(n / sum(n) * 100, 1), "%)", sep = "")),  # Add percentages
            position = position_stack(vjust = 0.5), size = 6, color = "white") +     # Add labels
  theme(
    plot.title = element_text(size = 14, hjust = 0),
    legend.text = element_text(size = 14),  # Set legend size
    legend.title = element_text(size = 14)  
  )

# Save plot as PNG
png("plots/evidence_types.png", width = 350, height = 300)
print(evidence_types)
dev.off()
```

# Clean the data

```{r}
# Remove NAs
if (nrow(gutmgene%>%filter(is.na(Metabolite_CID))) != 0){
  cat("Removing", nrow(gutmgene%>%filter(is.na(Metabolite_CID))), "Metabolite_CID NAs\n")
  gutmgene_na <- gutmgene %>%
    filter(!is.na(Metabolite_CID))
} else{
  cat("There are", nrow(gutmgene%>%filter(is.na(Metabolite_CID))), "Metabolite_CID NAs\n")
}

cat("N metabolites before removing duplicates:", length(gutmgene$Metabolite_CID), "\n")
# Remove duplicates
gutmgene_clean <- gutmgene_na %>%
  distinct(Metabolite_CID, .keep_all=TRUE)
cat("N metabolites without duplicates:", length(gutmgene_clean$Metabolite_CID), "\n")
```

# Get SMILES

```{r}
# Run the get_smiles_from_cid() for each metabolite CID
smiles <- sapply(gutmgene_clean$Metabolite_CID, get_smiles_from_cid)
cat(length(smiles), "SMILES were obtained\n")

# Create dataframe to store SMILES and CIDs for target prediction
df_smiles_ids <- data.frame(SMILES=smiles,
                         pubchem_ids=paste0("cid", gutmgene_clean$Metabolite_CID),
                         stringsAsFactors=FALSE)

# Save the dataframe as text file
write.table(df_smiles_ids, file="smiles_ids.txt",
            row.names=FALSE, col.names=FALSE,
            quote=FALSE, sep=" ")
```

# Parsing SEA results

After running SEA with smiles_ids.txt, read the results.

```{r}
# Read SEA dataframe
sea_result <- read.csv("data/sea-results.xls")

# Rename columns to avoid special character issues
colnames(sea_result) <- c("Query_ID", "Target_ID", "Affinity_Threshold", "P_Value",
                          "Max_Tc", "Cut_Sum", "Z_Score", "Name", "Description", "Query_Smiles")

# Get only the human targets 
sea_result <- sea_result[grepl("_HUMAN$", sea_result$Target_ID), ]

# Remove duplicates
sea_targets <- sea_result[!duplicated(sea_result$Target_ID), ]
cat("There are", length(sea_targets$Target_ID), "targets predicted with SEA\n")


# Filter for p <= 0.01
p_value <- 0.01
sea_targets_filtered <- sea_targets[sea_targets$P_Value <= p_value, ]
cat(length(sea_targets_filtered$Target_ID), "targets with SEA with p <= 0.01\n")
```

# Parsing STP results

