<!-- # --- -->
title: "Untitled"
author: "vitoria"
date: "2025-04-16"
output: html_document
---

```{r message=FALSE, warning=FALSE, include=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

# Pacotes do CRAN necessários
cran_pkgs <- c(
  "data.table", "dplyr", "future", "future.apply", "ggplot2", "rlang", "irlba", 
  "tidyr", "patchwork", "stringi", "fastmatch", "RcppRoll", "scales", "Rcpp", 
  "devtools",
  "tidyselect", "vctrs", "lifecycle", "dotCall64", "RcppEigen", "png", "reticulate"
)

# Pacotes do Bioconductor necessários
bioc_pkgs <- c(
  "GenomeInfoDb", "GenomicRanges", "IRanges", "Rsamtools", "S4Vectors", 
  "SeuratObject", "Biostrings", "XVector", "Rhtslib", "BiocParallel", 
  "UCSC.utils", "GENIE3", "variancePartition"
)

# Função pra instalar se faltar
install_missing <- function(pkgs, installer) {
  for (pkg in pkgs) {
    if (!requireNamespace(pkg, quietly = TRUE)) {
      message(paste("Instalando:", pkg))
      installer(pkg)
    } else {
      message(paste("Já está instalado:", pkg))
    }
  }
}

# Instala os pacotes
install_missing(cran_pkgs, install.packages)
install_missing(bioc_pkgs, BiocManager::install)

if (!requireNamespace("remotes", quietly = TRUE))
    install.packages("remotes")
remotes::install_github("stuart-lab/signac", ref = "develop")

devtools::install_github("immunogenomics/presto")
```

```{r message=FALSE, warning=FALSE}
# Load necessary libraries

# library(GenomeInfoDb)
# library(ggplot2)
# library(patchwork)
library(dplyr)
library(Seurat)
library(Matrix)
# library(GENIE3)
# library(variancePartition)
# library(BiocParallel)
# library(WGCNA)
```

# Functions

```{r}
#' Builds a Seurat object from snRNA-seq data
#'
#' @param matrix_file String - Path to a .mtx.gz file
#' @param genes_file String - Path to file with gene names
#' @param cells_file String - Path to file with cell names
#' @param project_name String - Name of your project/dataset
#' @param min.cells Integer - Minimum number of cells a gene must be detected in to be retained
#' @param min.features Integer - Minimum number of genes a cell must express to be included
#' @param nfeatures Integer - Number of top variable features to select.
#' @return seurat_obj - SeuratObject for downstream analyses
process_seurat_object <- function(matrix_file, genes_file, cells_file, group_file,
                                  project_name = "SeuratProject",
                                  min_cells = 3, min_features = 100) {
  
  # Load expression matrix
  expr_matrix <- readMM(matrix_file)
  
  # Load gene and cell names
  genes <- read.csv(genes_file, header = TRUE)
  cells <- read.csv(cells_file, header = TRUE)
  
  # Define row and column names
  rownames(expr_matrix) <- make.unique(genes$x)
  colnames(expr_matrix) <- cells$x
  
  # Create object
  seurat_obj <- CreateSeuratObject(counts = expr_matrix, 
                                   project = project_name, 
                                   min.cells = min_cells, 
                                   min.features = min_features)
  
  return(seurat_obj)
}

#' Builds a Seurat object from snRNA-seq data
#'
#' @param seurat_obj SeuratObject
#' @param nfeatures Integer - Number of features to filter if your dataset is large (e.g. top 10000 genes)
#' @param group_by String - Name of the variable you want to group your data by (condition, treatment, etc)
#' @param ident_1 String - Name of the first group considering the condition above
#' @param ident_2 String - Name of the second group considering the condition above
#' @param method String - Statistical test for differential expression ("wilcox", "t", "bimod", "LR")
#' @param logfc_threshold Integer - Threshold for FoldChange filtering (e.g. > 2)
#' @param pval_adj_threshold Float - Threshold for adjusted p value (e.g. <= 0.01) 
#' @param save Boolean - If you want to save your DEGs as a CSV
#' @param output_path String - Path to save the CSVs
#' @param prefix String - Prefix for the CSV filenames
#' @return list with p-adjusted DEGs and filtered DEGs
run_DE_analysis <- function(seurat_obj,
                            nfeatures = 20000,
                            group_by = "condition",
                            ident_1 = "MDD",
                            ident_2 = "Control",
                            method = "wilcox",
                            logfc_threshold = 1,
                            pval_adj_threshold = 0.05,
                            save = FALSE,
                            output_path = "",
                            prefix = "") {
  
  # Find DEGs
  degs <- FindMarkers(seurat_obj,
                      ident.1 = ident_1,
                      ident.2 = ident_2,
                      group.by = group_by,
                      test.use = method)
  
  # Adjust p value
  degs$p_val_adj <- p.adjust(degs$p_val, method = "fdr")
  
  # Filter by p value and FC
  degs_filtered <- degs %>%
    filter(p_val_adj < pval_adj_threshold, abs(avg_log2FC) > logfc_threshold)

  # Save if requested
  if (save){
    dir.create(output_path, showWarnings = FALSE, recursive = TRUE)

    write.csv(degs,
              file = file.path(output_path, paste0(prefix, "_results_adj.csv")),
              row.names = TRUE)
  
    write.csv(degs_filtered,
              file = file.path(output_path, paste0(prefix, "_results_filtered.csv")),
              row.names = TRUE)    
  }  

  return(list(degs_all = degs, degs_filtered = degs_filtered))
}

```



```{r}
gse_path <- "data/snrnaseq/GSE213982/"

# Build the Seurat object
GSE213982 <- process_seurat_object(
  matrix_file = paste0(gse_path, "GSE213982_combined_counts_matrix.mtx.gz"),
  genes_file = paste0(gse_path,  "GSE213982_combined_counts_matrix_genes_rows.csv.gz"),
  cells_file = paste0(gse_path, "GSE213982_combined_counts_matrix_cells_columns.csv.gz"),
  project_name = "GSE213982_snRNA"
)

# Gather meta data
cell_info <- do.call(rbind, strsplit(colnames(GSE213982), "\\."))
GSE213982$sample_id <- cell_info[,1]
GSE213982$barcode <- cell_info[,2]
GSE213982$cell_type <- cell_info[,3]
GSE213982$cell_subtype <- cell_info[,4]

# Gather groups information
condition_info <- read.csv(paste0(gse_path, "/groups.csv"))
GSE213982$condition <- condition_info$Condition[match(GSE213982$sample_id,
                                                       condition_info$Sample)]
GSE213982$condition[GSE213982$condition == "Case"] <- "MDD"

# Normalization
GSE213982 <- NormalizeData(GSE213982, normalization.method = "LogNormalize")

save(GSE213982, file=paste0(gse_path, "GSE213982_counts.RData"))
```
# Analysis without considering cell types

```{r}
gse_path <- "data/snrnaseq/GSE213982/"

load(paste0(gse_path, "GSE213982_counts.RData"))

degs_GSE213982 <- run_DE_analysis(paste0(GSE213982, "degs/"), save = TRUE,
                                  output_path=gse_path, prefix="GSE213982")
GSE213982 <- NULL
gc()
```



## GSE144136

```{r}
GSE144136 <- process_seurat_object(
  matrix_file = paste0(gse_path, "GSE144136_GeneBarcodeMatrix_Annotated.mtx.gz"),
  genes_file = paste0(gse_path, "GSE144136_GeneNames.csv.gz"),
  cells_file = paste0(gse_path, "GSE144136_CellNames.csv.gz"),
  project_name = "GSE144136_snRNA"
)

cell_info <- do.call(rbind, strsplit(colnames(GSE144136), "\\_"))
GSE144136$cell_type <- cell_info[,1]

GSE144136@meta.data$condition <- ifelse(grepl("Control", rownames(GSE144136@meta.data)), 
                                    "Control", 
                                    ifelse(grepl("Suicide", rownames(GSE144136@meta.data)), 
                                           "Suicide", 
                                           NA))

GSE144136$condition[GSE144136$condition == "Suicide"] <- "MDD"

GSE144136@meta.data$cell_type <- dplyr::case_when(
  GSE144136@meta.data$cell_type == "Astros" ~ "Ast",
  GSE144136@meta.data$cell_type == "Inhib" ~ "InN",
  GSE144136@meta.data$cell_type == "Ex" ~ "ExN",
  GSE144136@meta.data$cell_type == "Oligos" ~ "Oli",
  GSE144136@meta.data$cell_type == "OPCs" ~ "OPC",
  grepl("^Micro/Macro", GSE144136@meta.data$cell_type) ~ "Mic",
  grepl("^Endo", GSE144136@meta.data$cell_type) ~ "End",

  TRUE ~ GSE144136@meta.data$cell_type  
)

GSE144136 <- NormalizeData(GSE144136, normalization.method = "LogNormalize")

save(GSE144136, file=paste0(gse_path, "GSE144136_counts.RData"))
```


```{r}
gse_path <- "data/snrnaseq/GSE144136/"

load(paste0(gse_path, "GSE144136_counts.RData"))

degs_GSE144136 <- run_DE_analysis(paste0(GSE144136, "degs/"), save = TRUE,
                                  output_path=gse_path, prefix="/GSE144136")
GSE144136 <- NULL
gc()
```

## Getting the intersection between snRNA-seq and target prediction


```{r}
gse_path <- "data/snrnaseq/GSE144136/"
degs_GSE144136 <- read.csv(paste0(gse_path, "degs/GSE144136_results_filtered.csv"))

gse_path <- "data/snrnaseq/GSE213982/"
degs_GSE213982 <- read.csv(paste0(gse_path, "degs/GSE213982_results_filtered.csv"))

# Load target_intersection 
load(file='data/target_prediction/targets_intersection.RData')

snrnaseq_intersection <- intersect(degs_GSE213982, degs_GSE144136)
notype_intersection  <- intersect(snrnaseq_intersection, targets_intersection)

sea_targets <- read.csv("data/target_prediction/sea/sea_targets.csv")$x
stp_targets <- read.csv("data/target_prediction/stpnovo/stp_targets.csv")$x
all_targets <- union(sea_targets, stp_targets)
notype_intersection_alltargets <- intersect(snrnaseq_intersection, all_targets)

N = 22500 # Human genes
K = length(targets_intersection) # Intersection of SEA and STP predicted targets
n = length(snrnaseq_intersection) # DEGs found in the snRNAseq studies
k = length(notype_intersection_alltargets) # Intersection of TWAS and predicted targets
overlap_hypergeometric <- phyper(k-1, K, N - K, n, lower.tail = FALSE)

save(notype_intersection, notype_intersection_alltargets, file = "data/snrnaseq/degs_intersections_notype.RData")
```


# Analysis considering cell types

```{r message=FALSE, warning=FALSE, include=FALSE}
gse_path <- "data/snrnaseq/GSE213982/"
load(paste0(gse_path, "GSE213982_counts.RData"))

GSE213982_markers <- list()
cell_types <- setdiff(unique(GSE213982@meta.data$cell_type), "Mix")

for (ct in cell_types) {
  seurat_sub <- subset(GSE213982,
                       subset = cell_type == ct)
  
  seurat_sub <- NormalizeData(seurat_sub,
                              normalization.method = "LogNormalize")
  Idents(seurat_sub) <- "condition"

  markers <- FindMarkers(seurat_sub, ident.1 = "MDD",
                         ident.2 = "Control",
                         test.use = "wilcox")
  
  GSE213982_markers[[ct]] <- markers
  gc()
}

# filter by p and FC
GSE213982_markers_filtered <- list()
for (ct in cell_types) {
  markers <- GSE213982_markers[[ct]]
  
  markers_filtered <- markers %>%
    filter(p_val_adj < 0.05, abs(avg_log2FC) >= 1)

  GSE213982_markers_filtered[[ct]] <- markers_filtered
}

save(GSE213982_markers_filtered,
     file=paste0(gse_path, "type_degs/GSE213982_degs_type.RData"))
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
gse_path <- "data/snrnaseq/GSE144136/"
load(paste0(gse_path, "/GSE144136_counts.RData"))
# 
# GSE144136 <- FindVariableFeatures(GSE144136,
#                                      selection.method = "vst",
#                                      nfeatures = 1000)

GSE144136_markers <- list()
cell_types <- setdiff(unique(GSE144136@meta.data$cell_type), "Mix")

for (ct in cell_types) {
  seurat_sub <- subset(GSE144136, subset = cell_type == ct)
  seurat_sub <- NormalizeData(seurat_sub, normalization.method = "LogNormalize")

  Idents(seurat_sub) <- "condition"

  markers <- FindMarkers(seurat_sub, ident.1 = "MDD", ident.2 = "Control",
                         test.use = "wilcox")

  GSE144136_markers[[ct]] <- markers
}

# filter by p and FC
GSE144136_markers_filtered <- list()
for (ct in cell_types) {
  markers <- GSE144136_markers[[ct]]
  
  markers_filtered <- markers %>%
    filter(p_val_adj < 0.05, abs(avg_log2FC) >= 1)

  GSE144136_markers_filtered[[ct]] <- markers_filtered
}

save(GSE144136_markers_filtered,
     file=paste0(gse_path, "type_degs/GSE144136_degs_type.RData"))
```

## Getting the intersection between snRNA-seq and target prediction

```{r message=FALSE, warning=FALSE}
# Load target_intersection (303 targets)
load(file='data/target_prediction/targets_intersection.RData')

cell_types <- names(GSE144136_markers_filtered)

intersection_celltypes <- list()
for (ct in cell_types) {
  intersection_celltypes[[ct]] <- intersect(rownames(GSE144136_markers_filtered[[ct]]),
                                            rownames(GSE213982_markers_filtered[[ct]]))
}

type_intersection <- list()
for (ct in cell_types) {
  type_intersection[[ct]] <- intersect(targets_intersection,
                                               intersection_celltypes[[ct]])
}

sea_targets <- read.csv("data/target_prediction/sea/sea_targets.csv")$x
stp_targets <- read.csv("data/target_prediction/stpnovo/stp_targets.csv")$x
all_targets <- union(sea_targets, stp_targets)

type_intersection_alltargets <- list()
for (ct in cell_types) {
  type_intersection_alltargets[[ct]] <- intersect(all_targets,
                                               intersection_celltypes[[ct]])
}

save(type_intersection, type_intersection_alltargets, file = "data/snrnaseq/degs_intersections_type.RData")
```

# Comparing cell types

### Type by type

```{r message=FALSE, warning=FALSE, include=FALSE}
gene <- "NR4A1"
nr4a1_expressions <- list()
cell_types <- setdiff(unique(GSE213982@meta.data$cell_type), "Mix")
for (ct in cell_types) {
  
  gene_data <- FetchData(GSE213982, vars = c("NR4A1", "condition", "cell_type"))
  gene_data_filtered <- gene_data[gene_data$cell_type == ct, ]

  # Create violin plot
  nr4a1_condition_exp <- ggplot(gene_data_filtered, aes(x = condition,
                                                     y = NR4A1)) +
                          geom_boxplot(alpha = 0.7, outlier.alpha = 0.5) +
                          geom_jitter(width = 0.2, alpha = 0.3, size = 0.5) +
                          # geom_violin(alpha = 0.7, trim = FALSE) +
                          # geom_boxplot(width = 0.1, fill = "white",
                          #              alpha = 0.8) +
                          theme_light() +
                          labs(title = paste0(gene, " expression - ", ct),
                               x = "Condition", 
                               y = "NR4A1 Expression") +
                          theme(legend.position = "none",
                                plot.title = element_text(hjust = 0.5,
                                                          size = 14,
                                                          face = "bold"))

  ggsave(paste0(gse_path, gene, "_", ct, ".png"), nr4a1_condition_exp)
}
```

# Variance partitioning

```{r message=FALSE, warning=FALSE}
library(Seurat)
library(variancePartition)
library(BiocParallel)

extract_top_variable_genes <- function(
  gse_path,
  use_variable_features = TRUE,
  nfeatures = 15000,
  chunk_size = 1000,
  n_cores = 4
) {
  # Derivar nome base do diretório
  base_name <- basename(normalizePath(gse_path))
  counts_file <- file.path(gse_path, paste0(base_name, "_counts.RData"))
  
  # Criar ambiente temporário e carregar objeto
  tmp_env <- new.env()
  load(counts_file, envir = tmp_env)
  
  # Pega o primeiro objeto do .RData
  seurat_object <- tmp_env[[ls(tmp_env)[1]]]
  rm(tmp_env)
  gc()
  
  seurat_matrix <- GetAssayData(seurat_object, assay = "RNA", slot = "data")

  # Seleciona genes variáveis, se indicado
  if (use_variable_features) {
    seurat_object <- FindVariableFeatures(seurat_object,
                                          selection.method = "vst",
                                          nfeatures = nfeatures)
    variable_genes <- VariableFeatures(seurat_object)

    seurat_matrix <- seurat_matrix[variable_genes, ]
  }

  # Preparar metadados
  meta_data <- seurat_object@meta.data
  # Se orig.ident não tiver variação, usar sample_id (se disponível)
  if (length(unique(meta_data$orig.ident)) <= 1 && "sample_id" %in% colnames(meta_data)) {
    message("orig.ident has only one level — replacing with sample_id.")
    meta_data$orig.ident <- meta_data$sample_id
  }
  meta_data$condition <- as.numeric(factor(meta_data$condition))
  meta_data$cell_type <- as.numeric(factor(meta_data$cell_type))

  seurat_object <- NULL
  gc()

  # Fórmula para particionar variância
  form <- ~ (1 | orig.ident) + condition + cell_type

  # Parâmetro de paralelismo
  param <- SnowParam(workers = n_cores, type = "SOCK", progressbar = FALSE)

  # Rodar por chunks
  gene_names <- rownames(seurat_matrix)
  n_genes <- length(gene_names)
  # varPart_list <- list()
  
  dir.create(file.path(gse_path, "variance"), showWarnings = FALSE, recursive = TRUE)


  for (i in seq(1, n_genes, by = chunk_size)) {
    cat("Processing genes", i, "to", min(i + chunk_size - 1, n_genes), "\n")
    gene_subset <- gene_names[i:min(i + chunk_size - 1, n_genes)]
    matrix_chunk <- seurat_matrix[gene_subset, ]

    varPart_chunk <- fitExtractVarPartModel(matrix_chunk, form, meta_data, BPPARAM = param)
    # varPart_list[[length(varPart_list) + 1]] <- varPart_chunk
    save(varPart_chunk, file=paste0(gse_path, "/variance/chunk_", i, "_", i+ chunk_size -1, ".RData"))

    rm(matrix_chunk)
    gc()
  }
  message("Todos os chunks salvos em: ", file.path(gse_path, "variance"))

}

combine_varPart_results <- function(gse_path, top_percentage) {
  variance_dir <- file.path(gse_path, "variance")
  files <- list.files(variance_dir, pattern = "^chunk.*\\.RData$", full.names = TRUE)
  
  # print(files)

  varPart_list <- lapply(files, function(f) {
    load(f)
    return(varPart_chunk)
  })

  varPart_combined <- do.call(rbind, varPart_list)
  varPart_combined$gene <- rownames(varPart_combined)

  threshold <- quantile(varPart_combined$condition, top_percentage)
  top_genes <- subset(varPart_combined, condition >= threshold)

  return(top_genes)
}
```

```{r message=FALSE, warning=FALSE}
gse_path <- "data/snrnaseq/GSE213982/"

extract_top_variable_genes(gse_path,use_variable_features = F,
                           chunk_size = 1000, n_cores = 4)
top_genes_GSE213982 <- combine_varPart_results(gse_path, top_percentage = 0.98)

save(top_genes_GSE213982, file = paste0(gse_path, "variance/top_variance_GSE213982.RData"))


gse_path <- "data/snrnaseq/GSE144136/"

extract_top_variable_genes(gse_path, use_variable_features = F,
                           chunk_size = 1000, n_cores = 4)

top_genes_GSE144136 <- combine_varPart_results("data/snrnaseq/GSE144136/",
                                               top_percentage = 0.98)

save(top_genes_GSE144136, file = paste0(gse_path, "variance/top_variance_GSE144136.RData"))


sea_targets <- read.csv("data/target_prediction/sea/sea_targets.csv")$x
stp_targets <- read.csv("data/target_prediction/stpnovo/stp_targets.csv")$x
targets_intersection <- intersect(stp_targets, sea_targets)

all_targets <- unique(c(sea_targets, stp_targets))

common_degs <- intersect(top_genes_GSE213982$gene, top_genes_GSE144136$gene)
variance_intersection <- intersect(common_degs, targets_intersection)
variance_intersection_alltargets <- intersect(common_degs, all_targets)


save(variance_intersection, variance_intersection_alltargets, file = "data/snrnaseq/variance_intersection.RData")
```


# Getting the predicted metabolites/ligands for the intersection genes

```{r}
source("get_metabolites.R")  

load("data/all_analyses_comparison.RData")
rnaseq_genes <- gene_analyses_df[!gene_analyses_df$analyses %in% c("twas", "variance"), ]$gene

targets_to_metabolites <- get_metabolites(rnaseq_genes)

for (target in names(targets_to_metabolites)){
  mets <- targets_to_metabolites[[target]]
  new_mets <- c()
  for (met in mets){
    name <- df3[df3$CID == met,]$name
    new_mets <- c(new_mets, name)
  }
  targets_to_metabolites[[target]] <- new_mets
}

metabolites_to_targets <- list()

for (gene in names(targets_to_metabolites)) {
  print(gene)
  for (metabolite in targets_to_metabolites[[gene]]) {
    print(metabolite)
    metabolites_to_targets[[metabolite]] <- c(metabolites_to_targets[[metabolite]], gene)
  }
}
metabolites_to_targets <- lapply(metabolites_to_targets, unique)

save(targets_to_metabolites, metabolites_to_targets, file="targets_metabolites.RData")
```