<!-- # --- -->
title: "Untitled"
author: "vitoria"
date: "2025-04-16"
output: html_document
---

```{r message=FALSE, warning=FALSE, include=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

# Pacotes do CRAN necessários
cran_pkgs <- c(
  "data.table", "dplyr", "future", "future.apply", "ggplot2", "rlang", "irlba", 
  "tidyr", "patchwork", "stringi", "fastmatch", "RcppRoll", "scales", "Rcpp", 
  "devtools",
  "tidyselect", "vctrs", "lifecycle", "dotCall64", "RcppEigen", "png", "reticulate"
)

# Pacotes do Bioconductor necessários
bioc_pkgs <- c(
  "GenomeInfoDb", "GenomicRanges", "IRanges", "Rsamtools", "S4Vectors", 
  "SeuratObject", "Biostrings", "glmGamPoi"
)

# Função pra instalar se faltar
install_missing <- function(pkgs, installer) {
  for (pkg in pkgs) {
    if (!requireNamespace(pkg, quietly = TRUE)) {
      message(paste("Instalando:", pkg))
      installer(pkg)
    } else {
      message(paste("Já está instalado:", pkg))
    }
  }
}

# Instala os pacotes
install_missing(cran_pkgs, install.packages)
install_missing(bioc_pkgs, BiocManager::install)

if (!requireNamespace("remotes", quietly = TRUE))
    install.packages("remotes")
remotes::install_github("stuart-lab/signac", ref = "develop")

devtools::install_github("immunogenomics/presto")
```

```{r message=FALSE, warning=FALSE}
# Load necessary libraries

# library(GenomeInfoDb)
# library(ggplot2)
# library(patchwork)
library(dplyr)
library(Seurat)
library(Matrix)
library(glmGamPoi)
# library(variancePartition)
# library(BiocParallel)

future::plan("sequential")
Sys.setenv(OMP_NUM_THREADS = "1", MKL_NUM_THREADS = "1")

```

# Functions

```{r}
#' Builds a Seurat object from snRNA-seq data
#'
#' @param matrix_file String - Path to a .mtx.gz file
#' @param genes_file String - Path to file with gene names
#' @param cells_file String - Path to file with cell names
#' @param project_name String - Name of your project/dataset
#' @param min.cells Integer - Minimum number of cells a gene must be detected in to be retained
#' @param min.features Integer - Minimum number of genes a cell must express to be included
#' @param nfeatures Integer - Number of top variable features to select.
#' @return seurat_obj - SeuratObject for downstream analyses
process_seurat_object <- function(matrix_file, genes_file, cells_file,
                                  project_name = "SeuratProject",
                                  min_cells = 3, min_features = 100) {
  
  # Load expression matrix
  expr_matrix <- readMM(matrix_file)
  
  # Load gene and cell names
  genes <- read.csv(genes_file, header = TRUE)
  cells <- read.csv(cells_file, header = TRUE)
  
  # Define row and column names
  rownames(expr_matrix) <- make.unique(genes$x)
  colnames(expr_matrix) <- cells$x
  
  # Create object
  seurat_obj <- CreateSeuratObject(counts = expr_matrix, 
                                   project = project_name, 
                                   min.cells = min_cells, 
                                   min.features = min_features)
  
  return(seurat_obj)
}

#' Builds a Seurat object from snRNA-seq data
#'
#' @param seurat_obj SeuratObject
#' @param group_by String - Name of the variable you want to group your data by (condition, treatment, etc)
#' @param ident_1 String - Name of the first group considering the condition above
#' @param ident_2 String - Name of the second group considering the condition above
#' @param method String - Statistical test for differential expression ("wilcox", "t", "bimod", "LR")
#' @param logfc_threshold Integer - Threshold for FoldChange filtering (e.g. > 2)
#' @param pval_adj_threshold Float - Threshold for adjusted p value (e.g. <= 0.01) 
#' @param save Boolean - If you want to save your DEGs as a CSV
#' @param output_path String - Path to save the CSVs
#' @param prefix String - Prefix for the CSV filenames
#' @return list with p-adjusted DEGs and filtered DEGs
run_DE_analysis <- function(seurat_obj,
                            group_by = "condition",
                            ident_1 = "MDD",
                            ident_2 = "Control",
                            method = "wilcox",
                            min_pct = 0.1,
                            max_cells = 10000,
                            save = FALSE,
                            output_path = "",
                            prefix = "") {
  
  # Find DEGs
  degs <- FindMarkers(seurat_obj,
                      ident.1 = ident_1,
                      ident.2 = ident_2,
                      group.by = group_by,
                      test.use = method,
                      min.pct = min_pct,
                      max.cells.per.ident = max_cells)
  
  # Adjust p value
  degs$p_val_adj <- p.adjust(degs$p_val, method = "fdr")

  # Save if requested
  if (save){
    dir.create(output_path, showWarnings = FALSE, recursive = TRUE)

    write.csv(degs,
              file = file.path(output_path, paste0(prefix, "_results_adj.csv")),
              row.names = TRUE)
    save(degs, file=file.path(output_path, paste0(prefix, "_results_adj.RData")))
  }  

  return(degs)
}
```


```{r}
gse_path <- "data/snrnaseq/GSE213982/"

# Build the Seurat object
GSE213982 <- process_seurat_object(
  matrix_file = paste0(gse_path, "GSE213982_combined_counts_matrix.mtx.gz"),
  genes_file = paste0(gse_path,  "GSE213982_combined_counts_matrix_genes_rows.csv.gz"),
  cells_file = paste0(gse_path, "GSE213982_combined_counts_matrix_cells_columns.csv.gz"),
  project_name = "GSE213982_snRNA"
)

# Gather meta data
cell_info <- do.call(rbind, strsplit(colnames(GSE213982), "\\."))
GSE213982$sample_id <- cell_info[,1]
GSE213982$barcode <- cell_info[,2]
GSE213982$cell_type <- cell_info[,3]

GSE213982$cell_subtype <- cell_info[,4]
GSE213982$cell_subtype <- sub("^([A-Za-z]+)([0-9]+).*", "\\1\\2", GSE213982$cell_subtype)

# Gather groups information
condition_info <- read.csv(paste0(gse_path, "/groups.csv"))
GSE213982$condition <- condition_info$Condition[match(GSE213982$sample_id,
                                                       condition_info$Sample)]
GSE213982$condition[GSE213982$condition == "Case"] <- "MDD"

# Remove noise, mitochondrial genes, ribosomal genes, and other contamination
qc_metrics <- GSE213982@meta.data
feature_cutoffs <- quantile(qc_metrics$nFeature_RNA, probs = c(0.02, 0.98))
count_cutoffs   <- quantile(qc_metrics$nCount_RNA,   probs = c(0.02, 0.98))

p1 <- VlnPlot(GSE213982, features = "nFeature_RNA", slot = "counts", pt.size = 0) +
  scale_y_continuous(breaks = pretty(c(0, max(GSE213982$nFeature_RNA)), n = 20))
p2 <- VlnPlot(GSE213982, features = "nCount_RNA", slot = "counts", pt.size = 0) +
  scale_y_continuous(breaks = pretty(c(0, max(GSE213982$nCount_RNA)), n = 20))
p_violin_GSE213982 <- (p1 | p2)  # junta os dois mantendo eixos separados

# Percentual de genes mitocondriais (MT-)
GSE213982[["percent.mt"]] <- PercentageFeatureSet(GSE213982, pattern = "^MT-")
# Percentual de genes ribossomais (RPS/RPL)
GSE213982[["percent.ribo"]] <- PercentageFeatureSet(GSE213982, pattern = "^RPS|^RPL")
# Percentual de genes de hemoglobina (HB, exceto pseudogenes HBP)
GSE213982[["percent.hb"]] <- PercentageFeatureSet(GSE213982, pattern = "^HB[^(P)]")

GSE213982 <- subset(
  GSE213982,
  subset = nFeature_RNA > feature_cutoffs[1] &
            nFeature_RNA < feature_cutoffs[2] &
            nCount_RNA   > count_cutoffs[1] &
            nCount_RNA   < count_cutoffs[2] &
            percent.mt < 5 & percent.ribo < 40 & percent.hb < 1
)

saveRDS(GSE213982, "prenormalization_213.rds")

# Normalization in batches
GSE213982 <- readRDS("prenormalization_213.rds")
gc()
Sys.setenv(OMP_NUM_THREADS = "1", MKL_NUM_THREADS = "1")
future::plan("sequential")
ids <- unique(GSE213982$sample_id)
save(ids, file="213_ids.RData")

objs <- lapply(ids, function(id){
  x <- subset(GSE213982, subset = sample_id == id)
  x <- SCTransform(x, method="glmGamPoi", vst.flavor="v2",
                   vars.to.regress=c("percent.mt","nCount_RNA"),
                   variable.features.n=2000,
                   return.only.var.genes=FALSE,
                   conserve.memory=TRUE, verbose=TRUE)
  saveRDS(x, paste0("sct_", id, ".rds")); rm(x); gc(); NULL
})
# Integrate
list_sct <- lapply(paste0("sct_", ids, ".rds"), readRDS)
features <- SelectIntegrationFeatures(object.list = list_sct, nfeatures = 2000)
list_sct <- PrepSCTIntegration(object.list = list_sct, anchor.features = features)
anchors <- FindIntegrationAnchors(object.list = list_sct, normalization.method="SCT",
                                  anchor.features=features)
GSE213982_sct <- IntegrateData(anchorset = anchors, normalization.method="SCT")

DefaultAssay(GSE213982_sct) <- "SCT"
Idents(GSE213982_sct) <- "condition"
levels(GSE213982_sct)

save(GSE213982_sct, file=paste0(gse_path, "GSE213982_sct_counts.RData"))
```

# Analysis without considering cell types

```{r}
gse_path <- "data/snrnaseq/GSE213982/"
load(paste0(gse_path, "GSE213982_sct_counts.RData"))

degs_GSE213982 <- run_DE_analysis(GSE213982, save = T,
                                  output_path = paste0(gse_path, "degs_sct/"),
                                  prefix = "GSE213982")
GSE213982 <- NULL
gc()
```


## GSE144136

```{r}
gse_path <- "data/snrnaseq/GSE144136/"
GSE144136 <- process_seurat_object(
  matrix_file = paste0(gse_path, "GSE144136_GeneBarcodeMatrix_Annotated.mtx.gz"),
  genes_file = paste0(gse_path, "GSE144136_GeneNames.csv.gz"),
  cells_file = paste0(gse_path, "GSE144136_CellNames.csv.gz"),
  project_name = "GSE144136_snRNA"
)

cell_info <- do.call(rbind, strsplit(colnames(GSE144136), "\\_"))
GSE144136@meta.data$cell_type <- cell_info[,1]
GSE144136@meta.data$cell_subtype <- sub("^([A-Za-z/]+)[_\\./]([0-9]+).*", "\\1\\2",
                                        rownames(GSE144136@meta.data))

GSE144136@meta.data$condition <- ifelse(grepl("Control", rownames(GSE144136@meta.data)), 
                                    "Control", 
                                    ifelse(grepl("Suicide", rownames(GSE144136@meta.data)), 
                                           "MDD", 
                                           NA))

# Make it the same patter as the other dataset
## Cell type
GSE144136@meta.data$cell_type <- dplyr::case_when(
  GSE144136@meta.data$cell_type == "Astros" ~ "Ast",
  GSE144136@meta.data$cell_type == "Inhib" ~ "InN",
  GSE144136@meta.data$cell_type == "Ex" ~ "ExN",
  GSE144136@meta.data$cell_type == "Oligos" ~ "Oli",
  GSE144136@meta.data$cell_type == "OPCs" ~ "OPC",
  grepl("^Micro/Macro", GSE144136@meta.data$cell_type) ~ "Mic",
  grepl("^Endo", GSE144136@meta.data$cell_type) ~ "End",

  TRUE ~ GSE144136@meta.data$cell_type  
)
## Cell subtype
cell_subtype_info <- stringr::str_match(rownames(GSE144136@meta.data), "^([A-Za-z/]+)[_\\./]([0-9]+)")
cell_type <- cell_subtype_info[, 2]
subtype_num <- cell_subtype_info[, 3]
standardized_type <- dplyr::case_when(
  cell_type == "Astros" ~ "Ast",
  cell_type == "Inhib" ~ "InN",
  cell_type == "Ex" ~ "ExN",
  cell_type == "Oligos" ~ "Oli",
  cell_type == "OPCs" ~ "OPC",
  cell_type == "Endo" ~ "End",
  cell_type == "Micro/Macro" ~ "Mic",
  TRUE ~ cell_type
)
GSE144136@meta.data$cell_subtype <- paste0(standardized_type, subtype_num)

# Remove noise, mitochondrial genes, ribosomal genes, and other contamination
qc_metrics <- GSE144136@meta.data
feature_cutoffs <- quantile(qc_metrics$nFeature_RNA, probs = c(0.05, 0.95))
count_cutoffs   <- quantile(qc_metrics$nCount_RNA,   probs = c(0.05, 0.95))

p1 <- VlnPlot(GSE144136, features = "nFeature_RNA", slot = "counts", pt.size = 0) +
  scale_y_continuous(breaks = pretty(c(0, max(GSE144136$nFeature_RNA)), n = 20))
p2 <- VlnPlot(GSE144136, features = "nCount_RNA", slot = "counts", pt.size = 0) +
  scale_y_continuous(breaks = pretty(c(0, max(GSE144136$nCount_RNA)), n = 20))
p_violin_GSE144136 <- (p1 | p2)  # junta os dois mantendo eixos separados

# Percentual de genes mitocondriais (MT-)
GSE144136[["percent.mt"]] <- PercentageFeatureSet(GSE144136, pattern = "^MT-")
# Percentual de genes ribossomais (RPS/RPL)
GSE144136[["percent.ribo"]] <- PercentageFeatureSet(GSE144136, pattern = "^RPS|^RPL")
# Percentual de genes de hemoglobina (HB, exceto pseudogenes HBP)
GSE144136[["percent.hb"]] <- PercentageFeatureSet(GSE144136, pattern = "^HB[^(P)]")

GSE144136 <- subset(
  GSE144136,
  subset = nFeature_RNA > feature_cutoffs[1] &
            nFeature_RNA < feature_cutoffs[2] &
            nCount_RNA   > count_cutoffs[1] &
            nCount_RNA   < count_cutoffs[2] &
            percent.mt < 5 & percent.ribo < 40 & percent.hb < 1
)

saveRDS(GSE144136, "pre_144_real.rds")
GSE144136 <- readRDS( "pre_144_real.rds")

future::plan("sequential")
Sys.setenv(OMP_NUM_THREADS = "1", MKL_NUM_THREADS = "1")

# Normalization
GSE144136 <- SCTransform(GSE144136,
  method="glmGamPoi", vst.flavor="v2",
  vars.to.regress=c("percent.mt","nCount_RNA"),
  variable.features.n=2000,
  return.only.var.genes=TRUE,
  conserve.memory=TRUE, verbose=TRUE
)

DefaultAssay(GSE144136) <- "SCT"
Idents(GSE144136) <- "condition"
levels(GSE144136)

save(GSE144136, file=paste0(gse_path, "GSE144136_sct_counts.RData"))
```


```{r}
gse_path <- "data/snrnaseq/GSE144136/"
load(paste0(gse_path, "GSE144136_counts.RData"))

degs_GSE144136 <- run_DE_analysis(GSE144136, save = T,
                                  output_path = paste0(gse_path, "degs_sct/"),
                                  prefix = "/GSE144136")
GSE144136 <- NULL
gc()
```

## Getting the intersection between snRNA-seq and target prediction


```{r}
pval_adj_threshold <- 0.05
logfc_threshold <- 0.6

gse_path <- "data/snrnaseq/GSE144136/"
degs_GSE144136_adj <- read.csv(paste0(gse_path, "degs/GSE144136_results_adj.csv"))
degs_GSE144136_f <- degs_GSE144136_adj %>%
  filter(p_val_adj <= pval_adj_threshold, abs(avg_log2FC) >= logfc_threshold)

gse_path <- "data/snrnaseq/GSE213982/"
degs_GSE213982_adj <- read.csv(paste0(gse_path, "degs/GSE213982_results_adj.csv"))
degs_GSE213982_f <- degs_GSE213982_adj %>%
  filter(p_val_adj <= pval_adj_threshold, abs(avg_log2FC) >= logfc_threshold)

degs_GSE144136 <- degs_GSE144136_f$X
degs_GSE213982 <- degs_GSE213982_f$X

snrnaseq_intersection <- intersect(degs_GSE213982_f$X, degs_GSE144136_f$X)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Pasta com os CSVs das interseções
csv_folder <- "data/target_prediction/intersections"

# Lista fixa para intersectar
snrnaseq_intersection <- as.character(snrnaseq_intersection)  # garantir vetor de caracteres

# Lista para guardar os resultados
results_logn <- list()

# Ler cada arquivo CSV e fazer interseção com snrnaseq_intersection
csv_files <- list.files(csv_folder, pattern = "\\.csv$", full.names = TRUE)

for (file in csv_files) {
  # Read raw lines first
  raw_lines <- readLines(file, warn = FALSE)

  # Skip if all lines are either empty or just '""'
  if (all(trimws(raw_lines) %in% c("", "\"\""))) {
    message(paste("Skipping file with only '\"\"' or blank lines:", file))
    next
  }
  
  # Nome base do arquivo para usar como key
  key <- gsub("^intersection_|\\.csv$", "", basename(file))
  
  # Ler CSV
  # print(file)
  df <- read.csv(file, stringsAsFactors = FALSE)
  # print(file)
  # Garantir vetor
  vec <- as.character(df$target)
  
  # Interseção com snrnaseq_intersection
  inter <- intersect(vec, snrnaseq_intersection)
  
  # Salvar na lista
  results_logn[[key]] <- inter
}

save(results_logn, file="data/snrnaseq/results_logn.RData")

# all_union <- unique((c(all_sea, all_stp, all_chembl, all_xfp, all_ecfp)))
```

# Analysis considering cell types

```{r message=FALSE, warning=FALSE, include=FALSE}
gse_path <- "data/snrnaseq/GSE213982/"
load(paste0(gse_path, "GSE213982_counts.RData"))

GSE213982_celltype <- list()
cell_types <- setdiff(unique(GSE213982@meta.data$cell_type), "Mix")

for (ct in cell_types) {
  seurat_sub <- subset(GSE213982,
                       subset = cell_type == ct)
  
  seurat_sub <- NormalizeData(seurat_sub,
                              normalization.method = "LogNormalize")
  Idents(seurat_sub) <- "condition"

  markers <- FindMarkers(seurat_sub, ident.1 = "MDD",
                         ident.2 = "Control",
                         test.use = "wilcox")
  
  GSE213982_celltype[[ct]] <- markers
  gc()
}

save(GSE213982_celltype,
     file=paste0(gse_path, "type_degs/GSE213982_celltype_notfiltered.RData"))

# filter by p and FC
GSE213982_celltype_filtered <- list()
for (ct in cell_types) {
  markers <- GSE213982_celltype[[ct]]
  
  markers_filtered <- markers %>%
    filter(p_val_adj <= 0.05, abs(avg_log2FC) >= 0.6)

  GSE213982_celltype_filtered[[ct]] <- markers_filtered
}

save(GSE213982_celltype_filtered,
     file=paste0(gse_path, "type_degs/GSE213982_celltype.RData"))
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
gse_path <- "data/snrnaseq/GSE144136/"
load(paste0(gse_path, "/GSE144136_counts.RData"))

GSE144136_celltype <- list()
cell_types <- setdiff(unique(GSE144136@meta.data$cell_type), "Mix")

for (ct in cell_types) {
  seurat_sub <- subset(GSE144136, subset = cell_type == ct)
  seurat_sub <- NormalizeData(seurat_sub, normalization.method = "LogNormalize")

  Idents(seurat_sub) <- "condition"

  markers <- FindMarkers(seurat_sub, ident.1 = "MDD", ident.2 = "Control",
                         test.use = "wilcox")

  GSE144136_celltype[[ct]] <- markers
}

save(GSE144136_celltype,
     file=paste0(gse_path, "type_degs/GSE144136_celltype_notfiltered.RData"))

# filter by p and FC
GSE144136_celltype_filtered <- list()
for (ct in cell_types) {
  markers <- GSE144136_celltype[[ct]]
  
  markers_filtered <- markers %>%
    filter(p_val_adj < 0.05, abs(avg_log2FC) >= 0.6)

  GSE144136_celltype_filtered[[ct]] <- markers_filtered
}

save(GSE144136_celltype_filtered,
     file=paste0(gse_path, "type_degs/GSE144136_celltype.RData"))
```

## Getting the intersection between snRNA-seq and target prediction

```{r message=FALSE, warning=FALSE, include=FALSE}
# Pasta com os CSVs das interseções
csv_folder <- "data/target_prediction/intersections"
load("./data/snrnaseq/GSE144136/type_degs/GSE144136_celltype.RData")
load("./data/snrnaseq/GSE213982/type_degs/GSE213982_celltype.RData")

cell_types <- names(GSE144136_celltype_filtered)
intersection_celltypes <- list()
for (ct in cell_types) {
  intersection_celltypes[[ct]] <- intersect(rownames(GSE144136_celltype_filtered[[ct]]),
                                            rownames(GSE213982_celltype_filtered[[ct]]))
}

# Lista para guardar os resultados
results_logn_celltype <- list()

# Ler cada arquivo CSV e fazer interseção com snrnaseq_intersection
csv_files <- list.files(csv_folder, pattern = "\\.csv$", full.names = TRUE)

for (file in csv_files){
  for (ct in cell_types){
    
    # Prediction methods + cell type degs
    key <- paste0(basename(file), "_", ct)
    
    # Read prediction methods intersections
    df <- read.csv(file, stringsAsFactors = FALSE)
    vec <- as.character(df$target)
    
    # Intersect with DEGs
    inter <- intersect(vec, intersection_celltypes[[ct]])
    
    results_logn_celltype[[key]] <- inter
  }
}

save(results_logn_celltype, file = "data/snrnaseq/results_logn_celltype.RData")


# all_union <- unique((c(all_sea, all_stp, all_chembl, all_xfp, all_ecfp)))
```

# Analysis considering cell subtypes

```{r message=FALSE, warning=FALSE, include=FALSE}
gse_path <- "data/snrnaseq/GSE213982/"
load(paste0(gse_path, "GSE213982_counts.RData"))

GSE213982_markers_subcelltype <- list()
cell_subtypes <- setdiff(unique(GSE213982@meta.data$cell_subtype), "Mix")

for (cst in cell_subtypes) {
  seurat_sub <- subset(GSE213982,
                       subset = cell_subtype == cst)
  
  seurat_sub <- NormalizeData(seurat_sub,
                              normalization.method = "LogNormalize")
  Idents(seurat_sub) <- "condition"

  markers <- FindMarkers(seurat_sub, ident.1 = "MDD",
                         ident.2 = "Control",
                         test.use = "wilcox")
  
  GSE213982_markers_subcelltype[[cst]] <- markers
  gc()
}

# filter by p and FC
GSE213982_cellsubtype_filt <- list()
for (cst in names(GSE213982_markers_subcelltype)) {
  markers <- GSE213982_markers_subcelltype[[cst]]
  
  markers_filtered <- markers %>%
    filter(p_val_adj < 0.05, abs(avg_log2FC) >= 1)

  GSE213982_cellsubtype_filt[[cst]] <- markers_filtered
}

save(GSE213982_cellsubtype_filt,
     file=paste0(gse_path, "subtype_degs/GSE213982_subtype.RData"))
```

```{r message=FALSE, warning=FALSE, include=FALSE}
gse_path <- "data/snrnaseq/GSE144136/"
load(paste0(gse_path, "GSE144136_counts.RData"))

GSE144136_markers_subcelltype <- list()
cell_subtypes <- setdiff(unique(GSE144136@meta.data$cell_subtype), "Mix")

for (cst in cell_subtypes) {
  seurat_sub <- subset(GSE144136,
                       subset = cell_subtype == cst)
  
  if (!all(c("MDD", "Control") %in% unique(seurat_sub$condition))) {
    message("Skipping ", cst, " due to missing $condition(s)")
    next
  }
  
  seurat_sub <- NormalizeData(seurat_sub,
                              normalization.method = "LogNormalize")
  seurat_sub <- SetIdent(seurat_sub, value = "condition")

  markers <- FindMarkers(seurat_sub, ident.1 = "MDD",
                         ident.2 = "Control",
                         test.use = "wilcox")
  
  GSE144136_markers_subcelltype[[cst]] <- markers
  gc()
}

# filter by p and FC
GSE144136_cellsubtype_filt <- list()
for (cst in names(GSE144136_markers_subcelltype)) {
  markers <- GSE144136_markers_subcelltype[[cst]]
  
  markers_filtered <- markers %>%
    filter(p_val_adj < 0.05, abs(avg_log2FC) >= 1)

  GSE144136_cellsubtype_filt[[cst]] <- markers_filtered
}

save(GSE144136_cellsubtype_filt,
     file=paste0(gse_path, "subtype_degs/GSE144136_subtype.RData"))
```

## Getting the intersection between snRNA-seq and target prediction

```{r message=FALSE, warning=FALSE}
# Load target_intersection 
load(file='data/target_prediction/targets_intersection.RData')
load(file='data/snrnaseq/GSE144136/subtype_degs/GSE144136_subtype.RData')

cell_subtypes <- union(names(GSE144136_cellsubtype_filt), names(GSE213982_cellsubtype_filt))

intersection_cellsubtypes <- list()
for (cst in cell_subtypes) {CD81
  if (cst %in% names(GSE213982_cellsubtype_filt) &
      cst %in% names(GSE144136_cellsubtype_filt)) {
    intersection_cellsubtypes[[cst]] <- intersect(
      rownames(GSE144136_cellsubtype_filt[[cst]]),
      rownames(GSE213982_cellsubtype_filt[[cst]])
    )
  }
}

subtype_intersection <- list()
for (cst in cell_subtypes) {
  subtype_intersection[[cst]] <- intersect(targets_intersection,
                                               intersection_cellsubtypes[[cst]])
}

save(subtype_intersection, file = "data/snrnaseq/degs_intersections_subtype.RData")
```


# Getting the predicted metabolites/ligands for the intersection genes

```{r}
source("get_metabolites.R")  

all_targets <- read.table("data/targets.txt")$V1
selected_genes <- c("PDE4B", "PDE4D")

targets_to_metabolites_all <- get_metabolites(all_targets)
targets_to_metabolites_all <- lapply(targets_to_metabolites_all, unique)
targets_to_metabolites_all <- lapply(targets_to_metabolites_all, function(x) {
  gsub("cid", "", x, ignore.case = TRUE)
})
save(targets_to_metabolites_all, file="data/network/network_wcpg/targets_metabolites_all.RData")

targets_to_metabolites_PDE4B <- get_metabolites(selected_genes)
targets_to_metabolites_PDE4B <- lapply(targets_to_metabolites_PDE4B, unique)
save(targets_to_metabolites_PDE4B, file="data/network/targets_metabolites_PDE4B.RData")
```
