---
title: "vitoria-project"
author: "vitoria"
date: "2025-04-02"
output: html_document
---

# Load libraries

```{r message=FALSE, warning=FALSE}
# List of libraries to install
libraries <- c("dplyr", "httr2", "jsonlite", "ggplot2", "readxl", "rvest", 
               "readr", "stringr", "tidyr", "scales", "httr", "xml2")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
if (!require("biomaRt", quietly = TRUE))
  BiocManager::install("biomaRt")

# Function to install missing libraries
install_if_missing <- function(lib) {
  if (!require(lib, character.only = TRUE)) {
    install.packages(lib, dependencies = TRUE)
    library(lib, character.only = TRUE)
  }
}

# Apply the function to all libraries
sapply(libraries, install_if_missing)
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(httr2)
library(jsonlite)
library(ggplot2)
# library(readxl)
# library(rvest)
# library(readr)
library(stringr)
library(tidyr)
library(scales)
library(httr)
library(xml2)
library(biomaRt)
```

# Target prediction
## Functions to get SMILES from CID

```{r message=FALSE, warning=FALSE}
# Try to get CID from compound name
# If not possible, try to get CID from SID
# Param - name: compound name (character)
# Return - compound_cid: CID (integer)
get_cid <- function(name) {
  if (is.na(name) || name == "") return(NA)  # Handle missing names
  
  encoded_name <- URLencode(name, reserved = TRUE)
  
  # Try getting CID using compound name endpoint
  url_cid <- paste0("https://pubchem.ncbi.nlm.nih.gov/rest/pug/compound/name/", 
                    encoded_name, "/cids/JSON")
  response_cid <- tryCatch({
    req <- request(url_cid) |> req_perform()
    if (resp_status(req) == 200) {
      data <- resp_body_json(req)
      if (!is.null(data$IdentifierList$CID)) {
        compound_id <- data$IdentifierList$CID[[1]]
        returncompound_id  # Return first CID
      }
    }
    return(NA)
  }, error = function(e) return(NA))
  
  # If a CID was found, return it immediately
  if (!is.na(response_cid)) {
    return(response_cid)
  }
  
  # Pause to respect API rate limits
  Sys.sleep(0.5)  
  
  # Try getting SID using substance name endpoint
  url_sid <- paste0("https://pubchem.ncbi.nlm.nih.gov/rest/pug/substance/name/", encoded_name, "/sids/JSON")
  response_sid <- tryCatch({
    req <- request(url_sid) |> req_perform()
    if (resp_status(req) == 200) {
      data <- resp_body_json(req)
      if (!is.null(data$IdentifierList$SID) && length(data$IdentifierList$SID) > 0) {
        data$IdentifierList$SID[[1]]  # Store the SID 
      } else {
        NA
      }
    } else {
      NA
    }
  }, error = function(e) return(NA))
  
  # Pause to respect API rate limits
  Sys.sleep(0.5)  
  
  # Try getting CID using SID
  url_sid_cid <- paste0("https://pubchem.ncbi.nlm.nih.gov/rest/pug/substance/sid/", 
                        response_sid, "/JSON")
  response_cid <- tryCatch({
    req <- request(url_sid_cid) |> req_perform()
    if (resp_status(req) == 200) {
      data <- resp_body_json(req)
      if (!is.null(data$PC_Substances[[1]]$compound[[2]]$id$id$cid) && length(data$PC_Substances[[1]]$compound[[2]]$id$id$cid) > 0) {
        # print("Found CID using SID") 
        compound_id <- data$PC_Substances[[1]]$compound[[2]]$id$id$cid[[1]]
        return(compound_id)  # Return CID
      } 
    }
    return(NA)
  }, error = function(e) return(NA))
}

# Get SMILES from Compound ID
# Param - compound_id: CID (integer) 
# Return (character): SMILES for the compound
get_smiles_from_cid <- function(compound_id) {

  url <- paste0("https://pubchem.ncbi.nlm.nih.gov/rest/pug/compound/cid/", compound_id, "/property/CanonicalSMILES/JSON")
  
  response <- request(url) |> req_perform()
  
  # If there was a response, get the data
  if (resp_status(response) == 200) {
    data <- resp_body_json(response)
    if (!is.null(data$PropertyTable$Properties[[1]]$CanonicalSMILES)) {
      compound_smiles <- data$PropertyTable$Properties[[1]]$CanonicalSMILES
      return(compound_smiles)
    }
  }
  
  return(NA)
}
```

## Read and preprocess the data from gutMGene

```{r}
# Read the CSV file from gutMGene
gutmgene <- read.csv("data/gutmgene/gutmgene.csv", stringsAsFactors = FALSE)

# Rename columns to avoid special character issues
colnames(gutmgene) <- c("Host_Species", "Gut_Microbe_ID", "Rank",
                        "Metabolite_ID", "Evidence_Type", "Evidence_Amount")

# Extract the metabolite name from the "Metabolite_ID" column
# and microbe name from "Gut_Microbe_ID"
gutmgene <- gutmgene %>%
  mutate(
    Metabolite_Name = str_remove(Metabolite_ID, "\\([^()]*\\)$"), # Removes last set of parentheses and content
    Metabolite_Name = str_trim(Metabolite_Name), # Trims whitespace
    Gut_Microbe_Name = str_extract(Gut_Microbe_ID, "^[^(]+"), # Extracts text before '('
    Gut_Microbe_ID = as.integer(str_extract(Gut_Microbe_ID, "(?<=\\().*(?=\\))")), # Extracts numbers inside '()'
  )

# Remove a possibly incorrect microorganism (it's actually an insect)
gutmgene <- gutmgene[!grepl("Bacteria Latreille et al. 1825", gutmgene$Gut_Microbe_Name), ]
# Remove not identified species?
gutmgene <- gutmgene[!grepl("CEBAS", gutmgene$Gut_Microbe_Name), ]

# How many metabolites
cat("N metabolites (repeated):",
    length(gutmgene$Metabolite_ID), "\n")
cat("N metabolites (unique):",
    length(unique(gutmgene$Metabolite_ID)), "\n")

# Number of unique species
unique_species_count <- gutmgene %>% filter(Rank == 'species') %>%
  summarise(count = n_distinct(Gut_Microbe_ID)) %>% pull(count)
cat("N unique microbe IDs for species:", unique_species_count, "\n")

# Number of unique genera
gutmgene$Microbe_Genus <- sub(" .*", "", gutmgene$Gut_Microbe_Name)
gutmgene$Microbe_Genus <- gsub("\\[|\\]", "", gutmgene$Microbe_Genus)
cat("N unique genera:", length(unique(gutmgene$Microbe_Genus)), "\n")
```

## Build plots of raw data

### Plot 1 - Distribution of genera among metabolites

```{r}
# Count occurrences of each genus
genus_counts <- table(gutmgene$Microbe_Genus)

# Filter to only include genera with frequency > 5
genera_to_keep <- names(genus_counts[genus_counts >= 5])
gutmgene_filtered <- gutmgene[gutmgene$Microbe_Genus %in% genera_to_keep, ]

# Order from most frequent to less frequent
gutmgene_filtered_count <- gutmgene_filtered %>%
  count(Microbe_Genus) %>%
  mutate(Microbe_Genus = reorder(Microbe_Genus, -n))

title_portuguese <- "Distribuição dos microorganismos da microbiota intestinal"
x_axis_portuguese <- "Microorganismo (gênero)"
y_axis_portuguese <- "Número de metabólitos produzidos"

# Create the plot
microbe_distribution <- ggplot(gutmgene_filtered_count,
                               aes(x = Microbe_Genus, y = n)) +
  geom_bar(stat = "identity", fill = "gray30", width = 0.5) +
  theme_minimal() + 
  ggtitle(title_portuguese) +  
  theme(
    plot.title = element_text(size = 18, hjust = 0.5),  
    axis.title.y = element_text(size = 16),
    axis.title.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16, angle = 60, hjust = 1),
    panel.grid = element_blank(),  # Remove grid lines
    axis.line.x = element_line(color = "black", linewidth = 0.5),  # Only X-axis line
    axis.line.y = element_line(color = "black", linewidth = 0.5),  # Only Y-axis line
    axis.ticks = element_line(color = "black", linewidth = 1)      # Ensure ticks are visible
  ) +
  labs(x = x_axis_portuguese, 
       y = y_axis_portuguese) +
  scale_y_continuous(breaks = seq(0,max(gutmgene_filtered_count$n),
                                  by = 5))
# Create a directory for plots if it doesn't exist
if (!dir.exists("plots")) {
  dir.create("plots")
}

# Save plot as PNG
png("plots/gutmgene/microbe_distribution.png", width = 800, height = 500)
print(microbe_distribution)
dev.off()
```

## Clean the data

```{r}
# Parse Compound ID (CID)
# with vectorized function
get_cid_vec <- Vectorize(get_cid)
gutmgene <- gutmgene %>%
  mutate(
    Metabolite_CID = get_cid_vec(Metabolite_Name) # Obtains CID for the metabolite name
  )

# Remove NAs
if (nrow(gutmgene %>% filter(is.na(Metabolite_CID))) != 0) {
  gutmgene_na <- gutmgene %>% filter(is.na(Metabolite_CID))
  gutmgene_clean <- gutmgene %>% filter(!is.na(Metabolite_CID))
  cat("Removed", nrow(gutmgene_na), "Metabolite_CID NAs\n")
} else {
  gutmgene_clean <- gutmgene
  gutmgene_na <- data.frame()
  cat("No Metabolite_CID NAs found\n")
}

cat("N metabolites before removing duplicates:", length(gutmgene$Metabolite_CID), "\n")
# Remove duplicates
gutmgene_nodup <- gutmgene_clean %>%
  distinct(Metabolite_CID, .keep_all=TRUE)
cat("N metabolites without duplicates:", length(gutmgene_nodup$Metabolite_CID), "\n")

# Save processed dataframe
save(gutmgene_nodup, file = "data/gutmgene_nodup.RData")  # No duplicated metabolites
save(gutmgene_na, file = "data/gutmgene_na.RData")        # Only the NAs
```

## Get SMILES

```{r}
# Run the get_smiles_from_cid() for each metabolite CID
smiles <- sapply(gutmgene_nodup$Metabolite_CID, get_smiles_from_cid)
cat(length(smiles), "SMILES were obtained\n")

# Create dataframe to store SMILES and CIDs for target prediction
df_smiles_ids <- data.frame(SMILES=smiles,
                         pubchem_ids=paste0("cid", gutmgene_nodup$Metabolite_CID),
                         stringsAsFactors=FALSE)
# Save the dataframe as text file
write.table(df_smiles_ids, file="smiles_ids.txt",
            row.names=FALSE, col.names=FALSE,
            quote=FALSE, sep=" ")

# Create dataframe to store names and CIDs of the metabolites 
# that are not on PubChem to find the SMILES by the name
df_names_ids <- data.frame(name=gutmgene_na$Metabolite_Name,
                         pubchem_id=paste0("nocid", 1:length(gutmgene_na$Metabolite_Name)),
                         stringsAsFactors=FALSE)
# Save the dataframe as text file
write.table(df_names_ids, file="names_ids.txt",
            row.names=FALSE, col.names=FALSE,
            quote=FALSE, sep=" ")
```

## Parsing SEA results

After running SEA with smiles_ids.txt, read the results.

```{r}
# Read SEA dataframe
sea_result <- read.csv("data/sea/sea-results.xls")

# Rename columns to avoid special character issues
colnames(sea_result) <- c("Query_ID", "Target_ID", "Affinity_Threshold", "P_Value",
                          "Max_Tc", "Cut_Sum", "Z_Score", "Name", "Description", "Query_Smiles")

# Get only the human targets 
sea_result <- sea_result[grepl("_HUMAN$", sea_result$Target_ID), ]

# Sort the data frame by Target_ID and then by P_Value (ascending)
sea_result <- sea_result[order(sea_result$Target_ID, sea_result$P_Value), ]
# Keep only the first occurrence of each Target_ID (which now has the lowest P_Value)
sea_result <- sea_result[!duplicated(sea_result$Target_ID), ]
cat(length(sea_result$Target_ID), "targets were predicted with SEA\n")

sea_targets <- sea_result$Target_ID
sea_targets <- gsub("_HUMAN", "", sea_targets)

targets <- data.frame(Targets = sea_targets)
write.csv(targets, "data/sea/sea_targets.csv", row.names = FALSE)
```

## Parsing STP results

```{r}
# Directory with STP .csv files
directory <- "data/stp"

# Get the files
files <- list.files(path = directory, pattern = "\\.csv$", full.names = TRUE)

# Iterate over the files creating dataframes
dataframes <- list()
for (file_path in files) {
  if (file.info(file_path)$size > 0) {
    df <- read_csv(file_path, show_col_types = FALSE)
    dataframes[[length(dataframes) + 1]] <- df
  } else {
    cat("No targets predicted:", basename(file_path), "\n")
  }
}

# Concatenate dataframes
stp_result <- bind_rows(dataframes)

# Write final dataframe
write_csv(stp_result, file.path(directory, "stp-results.csv"))
```

```{r}
# Read STP dataframe
stp_result <- read.csv("data/stp/stp-results.csv")

# Rename columns to avoid special character issues
colnames(stp_result) <- c("Name", "Common_name", "Uniprot_ID",
                          "ChEMBL_ID", "Class",
                          "Probability", "Known_Actives")

# Remove targets with probability = 0
stp_result <- stp_result[stp_result$Probability != 0, ]

# Filter the rows with higher probability
stp_result <- stp_result[order(stp_result$Uniprot_ID, stp_result$Probability), ]

# Get unique targets
stp_result <- stp_result %>% distinct(Uniprot_ID, .keep_all=TRUE)
stp_targets <- stp_result$Uniprot_ID

cat(length(stp_result$Uniprot_ID), "targets were predicted with STP\n")
```

```{r}
# Connect to Ensembl
mart <- useMart("ensembl", dataset="hsapiens_gene_ensembl")

# Convert UniProt ID to HGNC symbol (gene name)
uniprot_to_symbol <- getBM(
  attributes=c("hgnc_symbol"),
  filters="uniprotswissprot",
  values=stp_targets,
  mart=mart
)

stp_targets_hgnc <- uniprot_to_symbol$hgnc_symbol

targets <- data.frame(Targets = stp_targets_hgnc)
write.csv(targets, "data/stp/stp_targets.csv", row.names = FALSE)

targets_intersection <- intersect(sea_targets, stp_targets_hgnc)
save(targets_intersection, file='data/targets_intersection.RData')
```

# Getting the predicted metabolites/ligands for the intersection genes

```{r}
sea_result <- read.csv("data/target_prediction/sea/sea-results.xls")

# Rename columns to avoid special character issues
colnames(sea_result) <- c("Query_ID", "Target_ID", "Affinity_Threshold", "P_Value",
                          "Max_Tc", "Cut_Sum", "Z_Score", "Name", "Description", "Query_Smiles")

# Get only the human targets 
sea_result <- sea_result[grepl("_HUMAN$", sea_result$Target_ID), ]

NR4A1 <- sea_result[sea_result$Name == "NR4A1", ]
metabolites_NR4A1B <- NR4A1$Query_ID

EDNRB <- sea_result[sea_result$Name == "EDNRB", ]
metabolites_EDNRB <- EDNRB$Query_ID

S1PR1 <- sea_result[sea_result$Name == "S1PR1", ]
metabolites_S1PR1 <- S1PR1$Query_ID
```
